# Predicting Salaries

## 1. Problem Definition

In this project we will be working with the UCI adult dataset. We will be attempting to predict if people in the data set belong in a certain class by salary, either making <=50k or >50k per year.

## 2. Data

http://archive.ics.uci.edu/ml/datasets/Adult

Attributes: 16

Instances: 32561

## 3. Evaluation

Binary classification - logistic regression - classification metrics

## 4. Features

age: continuous.

workclass: Private, Self-emp-not-inc, Self-emp-inc, Federal-gov, Local-gov, State-gov, Without-pay, Never-worked.

fnlwgt: continuous.

education: Bachelors, Some-college, 11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool.

education-num: continuous.

marital-status: Married-civ-spouse, Divorced, Never-married, Separated, Widowed, Married-spouse-absent, Married-AF-spouse.

occupation: Tech-support, Craft-repair, Other-service, Sales, Exec-managerial, Prof-specialty, Handlers-cleaners, Machine-op-inspct, Adm-clerical, Farming-fishing, Transport-moving, Priv-house-serv, Protective-serv, Armed-Forces.

relationship: Wife, Own-child, Husband, Not-in-family, Other-relative, Unmarried.

race: White, Asian-Pac-Islander, Amer-Indian-Eskimo, Other, Black.

sex: Female, Male.

capital-gain: continuous.

capital-loss: continuous.

hours-per-week: continuous.

country: United-States, Cambodia, England, Puerto-Rico, Canada, Germany, Outlying-US(Guam-USVI-etc), India, Japan, Greece, South, China, Cuba, Iran, Honduras, Philippines, Italy, Poland, Jamaica, Vietnam, Mexico, Portugal, Ireland, France, Dominican-Republic, Laos, Ecuador, Taiwan, Haiti, Columbia, Hungary, Guatemala, Nicaragua, Scotland, Thailand, Yugoslavia, El-Salvador, Trinadad&Tobago, Peru, Hong, Holand-Netherlands.

```{r}
df <- read.csv('adult_sal.csv')
head(df)
```
```{r}
str(df)
```

#####
## Categorical Features

```{r}
library(dplyr)
```

an attribute called `X` is a copy of index, let's drop this

```{r}
df <- select(df, -X)
```

There are 3 attributes that can be grouped into smaller groups

#####
**Type employer**

```{r}
table(df$type_employer)
```

```{r}
group_emp <- function(job){
  if(job == 'Local-gov' | job == 'State-gov'){
    return('SL-gov')
  } else if(job == 'Self-emp-inc' | job == 'Self-emp-not-inc'){
    return('Self-emp')
  } else{
    return(job)
  }
}
df$type_employer <- sapply(df$type_employer, group_emp)
```

```{r}
unemp <- function(job){
  if(job == 'Never-worked' | job == 'Without-pay'){
    return('Unemployed')
  } else{
    return(job)
  }
}
df$type_employer <- sapply(df$type_employer, unemp)
```

```{r}
table(df$type_employer)
```

######
**Marital status**

```{r}
table(df$marital)
```
```{r}
group_marital <- function(mar){
  mar <- as.character(mar)
  
  if(mar == 'Separated' | mar == 'Divorced' | mar == 'Widowed'){
    return('Not-married')
  }else if(mar == 'Never-married'){
    return(mar)
  }else{
    return('Married')
  }
}

df$marital <- sapply(df$marital, group_marital)
table(df$marital)
```

#####
**Country**

```{r}
table(df$country)
```
```{r}
Asia <- c('China','Hong','India','Iran','Cambodia','Japan', 'Laos' ,
          'Philippines' ,'Vietnam' ,'Taiwan', 'Thailand')

North.America <- c('Canada','United-States','Puerto-Rico' )

Europe <- c('England' ,'France', 'Germany' ,'Greece','Holand-Netherlands','Hungary',
            'Ireland','Italy','Poland','Portugal','Scotland','Yugoslavia')

Latin.and.South.America <- c('Columbia','Cuba','Dominican-Republic','Ecuador',
                             'El-Salvador','Guatemala','Haiti','Honduras',
                             'Mexico','Nicaragua','Outlying-US(Guam-USVI-etc)','Peru',
                            'Jamaica','Trinadad&Tobago')
Other <- c('South')
```

```{r}
group_country <- function(ctry){
  if(ctry %in% Asia){
    return('Asia')
  }else if(ctry %in% North.America){
    return('North-America')
  }else if(ctry %in% Europe){
    return('Europe')
  }else if(ctry %in% Latin.and.South.America){
    return('Latin-and-South-America')
  }else{
    return('Other')
  }
}
df$country <- sapply(df$country, group_country)
table(df$country)
```

Lastly, turn these categorical features into factors. Factor will convert the values into numerical categories.

```{r}
df$type_employer <- factor(df$type_employer)
df$education <- factor(df$education)
df$marital <- factor(df$marital)
df$occupation <- factor(df$occupation)
df$relationship <- factor(df$relationship)
df$race <- factor(df$race)
df$sex <- factor(df$sex)
df$country <- factor(df$country)
df$income <- factor(df$income)

str(df)
```

#####
## Dealing with Missing Data

```{r}
library(Amelia)
```

There are numbers of '?' values in `type_employer` and `occupation`. We need to turn these values into NA.

```{r}
df[df == '?'] <- NA
```

```{r}
table(df$type_employer)
```

Reassign the factor so '?' will not be included.

```{r}
df$type_employer <- factor(df$type_employer)
df$marital <- factor(df$marital)
df$occupation <- factor(df$occupation)
df$country <- factor(df$country)

str(df)
```
```{r}
missmap(df)
```

There are 1% missing values. We can drop these.

```{r}
df <- na.omit(df)
missmap(df)
```

## Some Visualization

```{r}
library(ggplot2)
```

```{r}
ggplot(df, aes(age)) + geom_histogram(aes(fill=income), color='black', binwidth=1) + theme_bw()
```
```{r}
ggplot(df, aes(hr_per_week)) + geom_histogram() + theme_bw()
```
```{r}
names(df)[names(df) == 'country'] <- 'region'

ggplot(df, aes(region)) + geom_bar(aes(fill=income), color='black') + theme_bw() + theme(axis.text.x = element_text(angle=90, hjust=1))
```

## Build the Model

```{r}
library(caTools)

set.seed(42)
sample <- sample.split(df$income, SplitRatio=0.7)
train <- subset(df, sample==TRUE)
test <- subset(df, sample==FALSE)
```

```{r}
model <- glm(income ~. , family=binomial(logit), data=train)
```
```{r}
summary(model)
```

It's quite confusing interpreting what these numbers mean. We can focus more on the P-value (Pr>|z|) and the significance level. In short, we can interpret these significance symbols (`*`) as how certain we can be that the coefficient has an impact on the dependent variable. Meaning that, the more (`*`), the more important the attribute.

## Predictions

```{r}
test$predicted.income <- predict(model, newdata=test, type='response')
```
```{r}
table(test$income, test$predicted.income > 0.5)
```
```{r}
# f1-score
(6372+1423)/(6372+1423+548+872)
```
```{r}
# recall
6732/(6372+548)
```
```{r}
# precision
6732/(6372+872)
```

## Conclusion

So what determines if people belong in the >50k per year?

* Age 40s
* Education Bachelors or above
* Married
* 40-60 hours per week

```{r}
ggplot(df, aes(marital)) + geom_bar(aes(fill=income), color='black') + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
```{r}
ggplot(df, aes(hr_per_week)) + geom_histogram(aes(fill=income), color='black') + theme_bw()
```
